---
title: "Conditions"
output:
  html_document:
    number_sections: yes
    theme: sandstone
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(tidy=TRUE, fig.align='center')
```

```{r, echo=FALSE, include=FALSE}
library(lobstr)
library(pryr)
library(knitr)
library(kableExtra)
library(magrittr)
library(dplyr)
library(tidyverse)
library(rlang)
```

# Conditions
## Introduction
The **condition** system provides a paired set of tools that allow the author of a function to indicate that something unusual is happening, and the user of that function to deal with it. The function author signals conditions with functions like `stop()` (for errors), `warning()` (for warnings), and `message()` (for messages), then the function user can handle them with functions like `tryCatch()` and `withCallingHandlers()`. Understanding the condition system is important because you’ll often need to play both roles: signalling conditions from the functions you create, and handle conditions signalled by the functions you call.

## Signalling conditions
There are three conditions that you can signal in code: errors, warnings, and messages.

* Errors are the most severe; they indicate that there is no way for a function to continue and execution must stop.

* Messages are the mildest; they are way of informing the user that some action has been performed on their behalf.

* Warnings fall somewhat in between, and typically indicate that something has gone wrong but the function has been able to at least partially recover.

There is a final condition that can only be generated interactively: an interrupt, which indicates that the user has “interrupted” execution by pressing Escape, Ctrl + Break, or Ctrl + C (depending on the platform).

```{r, error=TRUE}
stop("This is what an error looks like")
warning("This is what a warning looks like")
message("This is what a message looks like")
```

### Errors
In base R, errors are signalled, or thrown, by `stop()`:
```{r, error=TRUE}
f <- function() g()
g <- function() h()
h <- function() stop("This is an error!")

f()
```

### Warnings
Warnings, signalled by `warning()`, are weaker than errors: they signal that something has gone wrong, but the code has been able to recover and continue. Unlike errors, you can have multiple warnings from a single function call:
```{r, eval=FALSE}
fw <- function() {
  cat("1\n")
  warning("W1")
  cat("2\n")
  warning("W2")
  cat("3\n")
  warning("W3")
}

fw()
```

```
1
2
3
Warning messages:
1: In fw() : W1
2: In fw() : W2
3: In fw() : W3
```

By default, warnings are cached and printed only when control returns to the top level. You can control this behaviour with the warn option:

* To make warnings appear immediately, set options(warn = 1).

* To turn warnings into errors, set * options(warn = 2). This is usually the easiest way to debug a warning, as once it’s an error you can use tools like `traceback()` to find the source.

* Restore the default behaviour with options(warn = 0).

### Messages
`messages()` are displayed immediately and do not have a call. argument:
```{r}
fm <- function() {
  cat("1\n")
  message("M1")
  cat("2\n")
  message("M2")
  cat("3\n")
  message("M3")
}
```

Generally any function that produces a message should have some way to suppress it, like a `quiet = TRUE` argument. It is possible to suppress all messages with `suppressMessages()`.

### Exercises
1. Write a wrapper around `file.remove()` that throws an error if the file to be deleted does not exist.  
**A**:
    ```{r, error=TRUE}
    file_remove <- function(...) {
      if (!suppressWarnings(file.remove(...))) stop ("No such file")
    }
    
    file_remove("nosuchfile.no")
    ```

2. What does the appendLF argument to `message()` do? How is it related to `cat()`?  
**A**: indicates if a new line should be added to messages. `cat()` requires a `\n` for a new line.
    ```{r}
    message("Hello World!", appendLF = TRUE)
    message("Hello World!", appendLF = FALSE)
    message("Hello World\n", appendLF = FALSE)
    message("Hello World!", appendLF = TRUE)
    ```

3. What does `options(error = recover)` do? Why might you use it?  
**A**: When an error occurs, the standard R browser is then invoked from the corresponding environment; the user can type ordinary R language expressions to be evaluated in that environment.

4. What does `options(error = quote(dump.frames(to.file = TRUE)))` do? Why might you use it?  
**A**: Dumps error information to a file. Might be used for post-mortem debugging in non-interactive sessions.








signalConditon
q or quit
getermessage
debugger
browser
