---
title: "Debugging"
output:
  html_document:
    number_sections: yes
    theme: sandstone
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(tidy=TRUE, fig.align='center')
```

```{r, echo=FALSE, include=FALSE}
library(lobstr)
library(pryr)
library(knitr)
library(kableExtra)
library(magrittr)
library(dplyr)
library(tidyverse)
library(rlang)
```

# Debugging



debugger
dump.frames

browser
traceback
recover



browserText(n = 1)
browserCondition(n = 1)
browserSetDebug(n = 1

debug(fun, text = "", condition = NULL, signature = NULL)
debugonce(fun, text = "", condition = NULL, signature = NULL)
undebug(fun, signature = NULL)
isdebugged(fun, signature = NULL)
debuggingState(on = NULL)


debugCall
undebugcall
trace(what, tracer, exit, at, print, signature,
      where = topenv(parent.frame()), edit = FALSE)
untrace(what, signature = NULL, where = topenv(parent.frame()))
tracingState(on = NULL)
.doTrace(expr, msg)
returnValue(default = NULL)




findLineNum(srcfile, line, nameonly = TRUE,
            envir = parent.frame(), lastenv)

setBreakpoint(srcfile, line, nameonly = TRUE,
              envir = parent.frame(), lastenv, verbose = TRUE,
              tracer, print = FALSE, clear = FALSE, ...)


signature (S4 functions)

