---
title: "Functionals"
output:
  html_document:
    number_sections: yes
    theme: sandstone
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(tidy=TRUE, fig.align='center')
```

```{r, echo=FALSE, include=FALSE}
library(lobstr)
library(pryr)
library(knitr)
library(kableExtra)
library(magrittr)
library(dplyr)
library(tidyverse)
library(rlang)
```

R has what’s known as first class functions. You can do anything with functions that you can do with vectors: you can assign them to variables, store them in lists, pass them as arguments to other functions, create them inside functions, and even return them as the result of a function. A higher-order function is a function that takes a function as an input or returns a function as output. This part of the book is broken down by the three types of higher-order functions:

* **Functionals**, functions are the input. Functionals are by far and away the most immediately useful application of FP ideas, and you’ll use them all the time in data analyses.

* **Function factories**, functions as output. You can almost always avoid function factories in favour of a different technique, but they are occassionally useful.

* **Function operators**, functions as input and output. These are like adverbs, because they typically modify the operation of a function.

# Functionals
## Introduction
A *functional* is a function that takes a function as an input and returns a vector as output. Here’s a simple functional: it calls the function provided as input with 1000 random uniform numbers.

```{r}
randomise <- function(f) f(runif(1e3))
randomise(mean)
randomise(mean)
randomise(sum)
```

The three most frequently used are `lapply()`, `apply()`, and `tapply()`. All three take a function as input (among other things) and return a vector as output.

## `lapply()`
The simplest functional is `lapply()`, which you may already be familiar with. `lapply()` takes a function, applies it to each element in a list, and returns the results in the form of a list. `lapply()` is written in C for performance, but we can create a simple R implementation that does the same thing:

```{r, eval=FALSE}
lapply2 <- function(x, f, ...) {
  out <- vector("list", length(x))
  for (i in seq_along(x)) {
    out[[i]] <- f(x[[i]], ...)
  }
  out
}
```

## Looping patterns
It’s useful to remember that there are three basic ways to loop over a vector:

1. loop over the elements: `for (x in xs)`
2. loop over the numeric indices: `for (i in seq_along(xs))`
3. loop over the names: `for (nm in names(xs))`

The first form is usually not a good choice for a for loop because it leads to inefficient ways of saving output. With this form it’s very natural to save the output by extending a data structure, like in this example:

```{r, eval=FALSE}
xs <- runif(1e3)
res <- c()
for (x in xs) {
  # This is slow!
  res <- c(res, sqrt(x))
}
```

This is slow because each time you extend the vector, R has to copy all of the existing elements. Instead, it’s much better to create the space you’ll need for the output and then fill it in. This is easiest with the second form:

```{r, eval=FALSE}
res <- numeric(length(xs))
for (i in seq_along(xs)) {
  res[i] <- sqrt(xs[i])
}
```

Just as there are three basic ways to use a for loop, there are three basic ways to use `lapply()`:

1. `lapply(xs, function(x) {})`
2. `lapply(seq_along(xs), function(i) {})`
3. `lapply(names(xs), function(nm) {})`

Typically you’d use the first form because `lapply()` takes care of saving the output for you. However, if you need to know the position or name of the element you’re working with, you should use the second or third form. Both give you an element’s position (`i`, `nm`) and value (`xs[[i]]`, `xs[[nm]]`).

### Exercises
1. Why are the following two invocations of `lapply()` equivalent?

    ```{r, eval=FALSE}
    trims <- c(0, 0.1, 0.2, 0.5)
    x <- rcauchy(100)
    
    lapply(trims, function(trim) mean(x, trim = trim))
    lapply(trims, mean, x = x)
    ```
**A**: `x = x` is passed to `mean()`, and so `trims` is positionally matched to the `trim` argument of `mean()`.

2. The function below scales a vector so it falls in the range [0, 1]. How would you apply it to every column of a data frame? How would you apply it to every numeric column in a data frame?

    ```{r}
    scale01 <- function(x) {
      rng <- range(x, na.rm = TRUE)
      (x - rng[1]) / (rng[2] - rng[1])
    }
    ```
**A**: 
    ```{r}
    lapply(mtcars, scale01) %>% data.frame() %>% head %>% print(digits = 2)
    
    mt2 <- mtcars %>% rownames_to_column()
    lapply(mt2, function (x) if(is.numeric(x)) scale01(x) else x) %>% data.frame() %>% head %>% print(digits = 2)
    ```

3. Use both for loops and `lapply()` to fit linear models to the `mtcars` using the formulas stored in this list:

    ```{r}
    formulas <- list(
      mpg ~ disp,
      mpg ~ I(1 / disp),
      mpg ~ disp + wt,
      mpg ~ I(1 / disp) + wt
    )
    ```
**A**: 
    ```{r}
    lapply(formulas, lm, mtcars)
    
    res <- vector("list", length(formulas))
    for (i in seq_along(formulas)) {
      res[[i]] <- lm(formulas[[i]], mtcars)
    }
    res
    ```

4. Fit the model `mpg ~ disp` to each of the bootstrap replicates of `mtcars` in the list below by using a `for` loop and `lapply()`. Can you do it without an anonymous function?

    ```{r}
    bootstraps <- lapply(1:3, function(i) {
      rows <- sample(1:nrow(mtcars), rep = TRUE)
      mtcars[rows, ]
    })
    ```
**A**: 
    ```{r}
    lapply(bootstraps, function(x) lm(mpg ~ disp, x))
    lapply(bootstraps, lm, formula = mpg ~ disp)
    
    res <- vector("list", length(bootstraps))
    for (i in seq_along(bootstraps)) {
      res[[i]] <- lm(mpg ~ disp, bootstraps[[i]])
    }
    res    
    ```

5. For each model in the previous two exercises, extract R^2^ using the function below.

    ```{r}
    rsq <- function(mod) summary(mod)$r.squared
    
    mod1 <- lapply(formulas, lm, mtcars)
    mod2 <- lapply(bootstraps, lm, formula = mpg ~ disp)
    lapply(mod1, rsq)
    lapply(mod2, rsq)
    ```

## For loop functionals: friends of `lapply()`
### Vector output: `sapply()` and `vapply()`
`sapply()` and `vapply()` are very similar to `lapply()` except they simplify their output to produce an atomic vector. While `sapply()` guesses, `vapply()` takes an additional argument specifying the output type. `sapply()` is great for interactive use because it saves typing, but if you use it inside your functions you’ll get weird errors if you supply the wrong type of input. `vapply()` is more verbose, but gives more informative error messages and never fails silently. It is better suited for use inside other functions.

```
lapply(X, FUN, ...)
sapply(X, FUN, ..., simplify = TRUE, USE.NAMES = TRUE)
vapply(X, FUN, FUN.VALUE, ..., USE.NAMES = TRUE)
```

The following example illustrates these differences. When given a data frame, `sapply()` and `vapply()` return the same results. When given an empty list, `sapply()` returns another empty list instead of the more correct zero-length logical vector.

```{r}
sapply(mtcars, is.numeric)
vapply(mtcars, is.numeric, logical(1))
sapply(list(), is.numeric)
vapply(list(), is.numeric, logical(1))
```

If the function returns results of different types or lengths, `sapply()` will silently return a list, while `vapply()` will throw an error. 

```{r, error=TRUE}
df <- data.frame(x = 1:10, y = letters[1:10])
sapply(df, class)
vapply(df, class, character(1))

df2 <- data.frame(x = 1:10, y = Sys.time() + 1:10)
sapply(df2, class)
vapply(df2, class, character(1))
```

Simplification in `sapply()` is only attempted if `X` has length greater than zero and if the return values from all elements of `X` are all of the same (positive) length. If the common length is one the result is a vector, and if greater than one is a matrix with a column corresponding to each element of `X`.

Simplification is always done in `vapply()`. This function checks that all values of `FUN` are compatible with the `FUN.VALUE`, in that they must have the same length and type.

```{r}
sapply(list(1:2, 3:4), sqrt)
vapply(list(1:2, 3:4), sqrt, numeric(2))
```

`USE.NAMES` - logical; if `TRUE` and if `X` is character, use `X` as `names` for the result unless it had names already. 

```{r, error=TRUE}
randomise <- function(f) match.fun(f)(runif(1e3))
sapply(c("mean", "sum"), randomise)
vapply(c("mean", "sum"), randomise, numeric(1))
```

Note, `sapply(*, simplify = FALSE, USE.NAMES = FALSE)` is equivalent to `lapply(*)`.

`sapply()` is a thin wrapper around `lapply()` that transforms a list into a vector in the final step. `vapply()` is an implementation of `lapply()` that assigns results to a vector (or matrix) of appropriate type instead of as a list. The following code shows a pure R implementation of the essence of `sapply()` and `vapply()` (the real functions have better error handling and preserve names, among other things).

```{r, eval=FALSE}
sapply2 <- function(x, f, ...) {
  res <- lapply2(x, f, ...)
  simplify2array(res)
}

vapply2 <- function(x, f, f.value, ...) {
  out <- matrix(rep(f.value, length(x)), nrow = length(f.value))
  for (i in seq_along(x)) {
    res <- f(x[[i]], ...)
    stopifnot(
      length(res) == length(f.value),
      typeof(res) == typeof(f.value)
    )
    out[ ,i] <- res
  }
  out
}
```

Some more complicated examples are:

``` {r, error=TRUE}
lapply(list(1:2, 3:4), function(x) outer(x, x)) # list of 2 matrices
sapply(list(1:2, 3:4), function(x) outer(x, x), simplify = FALSE) # list of 2 matrices

(z <- sapply(list(1:2, 3:4), function(x) outer(x, x), simplify = TRUE)) # matrix
dim(z)
is.matrix(z)
is.array(z)

(z <- sapply(list(1:2, 3:4), function(x) outer(x, x), simplify = "array")) # 3d array
# NOTE sapply calls simplify2array() with higher = (simplify == "array"))
dim(z)
is.matrix(z)
is.array(z)

vapply(list(1:2, 3:4), function(x) outer(x, x), outer(1:2, 1:2)) #3d array

fv <- matrix(data = NA_real_, nrow = 2, ncol = 2, dimnames = list(c("u", "v"), c("x", "y")))
vapply(list(a = 1:2, b = 3:4), function(x) outer(x, x), fv) #named 3d array
```

### Replicate
`replicate()` is a wrapper for the common use of sapply for repeated evaluation of an expression (which will usually involve random number generation).

```{r}
replicate

set.seed(456)
replicate(5, mean(runif(2)))
set.seed(456)
sapply(integer(5), function(x) mean(runif(2)))
set.seed(NULL)
```

Be careful about using replicate with additional parameters:

```{r, error=TRUE}
foo <- function(x = 1, y = 2) c(x, y)

bar <- function(n, ...) replicate(n, foo(...))
bar(3, 4, 5)
# does not work: Because integer 3 = (0,0,0), and 0 is passed to foo each time  

# This works
bar <- function(n, x, y) replicate(n, foo(x = x, y = y))
bar(3, 4, 5)
```

### Multiple inputs
With `lapply()`, only one argument to the function varies; the others are fixed. This makes it poorly suited for some problems. For example, how would you find a weighted mean when you have two lists, one of observations and the other of weights?

```{r}
# Generate some sample data
xs <- replicate(2, runif(3), simplify = FALSE) 
xs
ws <- replicate(2, rpois(3, 5) + 1, simplify = FALSE)

# simplify = FALSE to stop output being matrix
```

It’s easy to use `lapply()` to compute the unweighted means:

```{r}
unlist(lapply(xs, mean))
```

But how could we supply the weights to `weighted.mean()`? `lapply(x, means, w)` won’t work because the additional arguments to `lapply()` are passed to every call. We could change looping forms:

```{r}
unlist(lapply(seq_along(xs), function(i) {
  weighted.mean(xs[[i]], ws[[i]])
}))
```

This works, but it’s a little clumsy. A cleaner alternative is to use `Map()`, a variant of `lapply()`, where all arguments can vary. This lets us write:

```{r}
unlist(Map(weighted.mean, xs, ws))
```

If some of the arguments should be fixed and constant, use an anonymous function:

```{r}
xs[[2]][[1]] <- NA
xs
Map(function(x, w) weighted.mean(x, w, na.rm = TRUE), xs, ws)
```

`Map()` is a simple wrapper to `mapply()`:

```{r}
Map
```

I prefer `Map()` because:

* It’s equivalent to `mapply` with `simplify = FALSE`, which is almost always what you want.

* Instead of using an anonymous function to provide constant inputs, mapply has the MoreArgs argument that takes a list of extra arguments that will be supplied, as is, to each call. This breaks R’s usual lazy evaluation semantics, and is inconsistent with other functions.

```{r}
mapply(weighted.mean, xs, ws, MoreArgs = list(na.rm = TRUE))
```

### Exercises
TO DO

## Vector fuctionals
Another way of thinking about functionals is as a set of general tools for altering, subsetting, and collapsing vectors. Every functional programming language has three tools for this: `Map()`, `Reduce()`, and `Filter()`. We’ve seen `Map()` already, and the following sections describe `Reduce()`, a powerful tool for extending two-argument functions, and `Filter()`, a member of an important class of functionals that work with predicates, functions that return a single `TRUE` or `FALSE`.

### `Reduce()`
`Reduce()` reduces a vector, `x`, to a single value by recursively calling a function, `f`, two arguments at a time. It combines the first two elements with `f`, then combines the result of that call with the third element, and so on. Calling `Reduce(f, 1:3)` is equivalent to `f(f(1, 2), 3)`. `Reduce` is also known as fold, because it folds together adjacent elements in the list.

The following two examples show what `Reduce` does with an infix and prefix function:

```{r}
Reduce(`+`, 1:3) # -> ((1 + 2) + 3)
Reduce(sum, 1:3) # -> sum(sum(1, 2), 3)
```

The essence of `Reduce()` can be described by a simple for loop:

```{r, eval=FALSE}
Reduce2 <- function(f, x) {
  out <- x[[1]]
  for(i in seq(2, length(x))) {
    out <- f(out, x[[i]])
  }
  out
}
```

The real `Reduce()` is more complicated because it includes arguments to control whether the values are reduced from the left or from the right (`right`), an optional initial value (`init`), and an option to output intermediate results (`accumulate`).

`educe()` is an elegant way of extending a function that works with two inputs into a function that can deal with any number of inputs. It’s useful for implementing many types of recursive operations, like merges and intersections. Imagine you have a list of numeric vectors, and you want to find the values that occur in every element:

```{r}
l <- replicate(5, sample(1:10, 15, replace = T), simplify = FALSE)
str(l)
```

You could do that by intersecting each element in turn:

```{r}
intersect(intersect(intersect(intersect(l[[1]], l[[2]]), l[[3]]), l[[4]]), l[[5]])
```

That’s hard to read. With `Reduce()`, the equivalent is:

```{r}
Reduce(intersect, l)
Reduce(intersect, l, accumulate = TRUE)
```

A more sophisticated example:

```{r}
Funcall <- function(f, ...) f(...)
## n-fold iterate of a function, functional style:
Iterate <- function(f, n = 1) function(y) Reduce(Funcall, rep.int(list(f), n), init = y, right = TRUE)
## Continued fraction approximation to the golden ratio:
Iterate(function(x) 1 + 1 / x, 30)(3)

# what's going on
Iterate(function(x) 1 + 1 / x, 2)(3)

f <- function(x) 1 + 1 / x
rep.int(list(f),2)

r <- Funcall(function (x) 1+1/x, 3)
Funcall(function (x) 1+1/x, r)
```


### Predicate Functionals
A predicate is a function that returns a single `TRUE` or `FALSE`. There are three useful predicate functionals in base R: `Filter()`, `Find()`, and `Position()`.

* `Filter()` selects only those elements which match the predicate.
* `Find()` returns the first element which matches the predicate (or the last element if right = TRUE).
* `Position()` returns the position of the first element that matches the predicate (or the last element if right = TRUE).



`Filter()` applies the unary predicate function `f` to each element of `x` and returns the subset of `x` for which this gives `TRUE`.

```{r, error=TRUE}
df <- data.frame(x = 1:3, y = c("a", "b", "c"), z = c("u", "v", "w"))

str(Filter(is.factor, df))
```

`Find(f, x, right = FALSE, nomatch = NULL)`  
`Position(f, x, right = FALSE, nomatch = NA_integer_)`

`Find()` and `Position()`: If there is an element for which the predicate function gives `TRUE`, then the first or last such element or its position is returned depending on whether right is `FALSE` (default) or `TRUE`, respectively. If there is no such element, the value specified by `nomatch` is returned.

```{r}
str(Find(is.factor, df))
Position(is.factor, df, right = TRUE)
```

Another useful predicate functional is `where()`, a custom functional that generates a logical vector from a list (or a data frame) and a predicate:

```{r}
where <- function(f, x) {
  vapply(x, f, logical(1))
}

where(is.factor, df)
```

`Negate(f)`: Given a (predicate) function `f`, it creates a function which returns the logical negation of what `f` returns.

```{r}
Negate
```


### Exercises
1. Why isn’t `is.na()` a predicate function? What base R function is closest to being a predicate version of `is.na()`?  
**A**: `is.na()` is a vectorised function. `AnyNA()` is a predicate version. 
    ```{r, error=TRUE}
    v <- c(1, NA, 3, NA)
    is.na(v)
    anyNA(v)
    
    df <- data.frame(x = v, y = c("a", "b", "c", "d"))
    Filter(is.na, df)
    Filter(anyNA, df)
    ```

2. Use `Filter()` and `vapply()` to create a function that applies a summary statistic to every numeric column in a data frame.
    ```{r, error=TRUE}
    df <- data.frame(x = 1:3, y = c("a", "b", "c"), z = 4:6)
    ss <- function(f, df) vapply(Filter(is.numeric, df), f, numeric(1L))
    ss(sum, df)
    ```

3. What’s the relationship between `which()` and `Position()`? What’s the relationship between `where()` and `Filter()`?  
**A**: `which()` returns all indices of true entries from a logical vector. `Position()` returns just the first (default) or the last integer index of all true entries that occur by applying a predicate function on a vector. 

    ```{r}
    v <- list(1, 2, "a")
    Position(is.numeric, v)
    min(which(vapply(v, is.numeric, logical(1L)))) 
    ```
    `Filter(f, x) <=> x[where(f, x)]`
    ```{r, error=TRUE}
    df <- data.frame(x = 1:3, y = c("a", "b", "c"), z = c("u", "v", "w"))
    
    Filter(is.factor, df)
    
    where(is.factor, df)
    df[where(is.factor, df)]
    ```

4. Implement `Any()`, a function that takes a list and a predicate function, and returns TRUE if the predicate function returns TRUE for any of the inputs. Implement `All()` similarly.
**A**: 
    ```{r}
    # Any
    u <- list("a", 2)
    v <- list("a", "b")
    Any <- function(f, x) any(vapply(x, f, logical(1L)))
    Any(is.numeric, u)
    Any(is.numeric, v)
    
    # All
    w <- list(1, 2)
    All <- function(f, x) all(vapply(x, f, logical(1L)))
    All(is.numeric, v)
    All(is.numeric, w)
    ```

5. Implement the `span()` function from Haskell: given a list `x` and a predicate function `f`, span returns the location of the longest sequential run of elements where the predicate is true. (Hint: you might find `rle()` helpful.)
**A**: 
    ```{r, error=TRUE}
    x <- c(1,1,3,3,3,3,2,2,1,1,1,1,4,4,5,6)
    x
    y <- rle(x)
    y
    z <- inverse.rle(y)
    z
    
    # span
    v <- list(1, 2, "a", "b", "c", "d", 3, 4, 5)
    w <- list(1, 2, 2, 2, "a", "b", "c", "d", 3, 4, 5)
    
    span <- function(f, x) {
      
      y <- rle(vapply(x, function(x) f(x), logical(1L)))
      z <- y$lengths
      z[!y$values] <- 0L
      
      p <- which.max(z) - 1L
      if (p == 0L) return (1L)
      cumsum(y$lengths)[p] + 1
    }
    
    span(is.numeric, v)
    span(is.numeric, w)
    ```

## Manipulating matrices and dataframes
### `apply()`
So far, all the functionals we’ve seen work with 1d input structures. The three functionals in this section provide useful tools for working with higher-dimensional data structures. `apply()` is a variant of `sapply()` that works with matrices and arrays. You can think of it as an operation that summarises a matrix or array by collapsing each row or column to a single number. It has four arguments:

* `X`, the matrix or array to summarise
* `MARGIN`, an integer vector giving the dimensions to summarise over, 1 = rows, 2 = columns, etc.
* `FUN`, a summary function
* `...` other arguments passed on to FUN

A typical example of `apply()` looks like this

```{r, error=TRUE}
m <- matrix(1:20, nrow = 5)
apply(m, 1, mean)
apply(m, 2, mean)

a <- array(1:20, c(2, 5, 2))
a
apply(a, c(1,2), sum)
```

There are a few caveats to using `apply()`. It doesn’t have a `simplify` argument, so you can never be completely sure what type of output you’ll get. This means that `apply()` is not safe to use inside a function unless you carefully check the inputs. 

If each call to `FUN` returns a vector of length `n`, then apply returns an array of dimension `c(n, dim(X)[MARGIN])` if `n > 1`. If `n` equals `1`, `apply` returns a vector if `MARGIN` has length `1` and an array of `dimension dim(X)[MARGIN]` otherwise. If `n` is `0`, the result has length `0` but not necessarily the ‘correct’ dimension.

If the calls to `FUN` return vectors of different lengths, apply returns a list of length `prod(dim(X)[MARGIN])` with `dim` set to `MARGIN` if this has length greater than one.

`apply()` is also not idempotent in the sense that if the summary function is the identity operator, the output is not always the same as the input:

```{r}
m <- matrix(1:20, nrow = 5)
dim(m)
m2 <- apply(m, 1, identity)
dim(m2)
m3 <- apply(m, 2, identity)
dim(m3)
```

### `sweep()`
`sweep()` applies the same operation to selected indices of an array,  
`sweep(x, MARGIN, STATS, FUN = "-", check.margin = TRUE, ...)`. `FUN` should be a function of two arguments: it will be called with arguments `x` and an array of the same dimensions generated from `STATS` by `aperm`.

```{r}
x <- matrix(1:20, nrow = 4)
sweep(x, 1, 4:1, `-`) # subract 4 from row 1, 3 from row 2, etc.
sweep(x, 1, 4, `-`) # subtract 4 from each row
```

### `split()`
`split()` divides the data in the vector `x` into the groups defined by `f`: `split(x, f, drop = FALSE)`

```{r, error=TRUE}
x <- 1:10
f <- factor(c(1, 1, 2, 3, 2, 3, 1, 2, 1, 3))
f

xf <- split(x, f)
xf

unsplit(xf, f)

# split into even and odd
split(1:10, 1:2)

# split a matrix into a list by columns
m <- matrix(1:20, nrow = 4)
col(m)
split(m, col(m))

# sum by split
y <- x
( split(y, f) <- lapply(split(y, f), sum) )

# can also spllit a dataframe
df <- data.frame(x = c(1,2,3), y = c(4,5,6), z = c(7,8,9))
f <- factor(c(1, 2, 1))
dff <- split(df, f)
dff

split.data.frame
```

### `tapply()`
Apply a function to each cell of a ragged array, that is to each (non-empty) group of values given by a unique combination of the levels of certain factors.

`tapply(X, INDEX, FUN = NULL, ..., default = NA, simplify = TRUE)`

```{r, error=TRUE}
x <- 1:10
f <- factor(c(1, 1, 2, 3, 2, 3, 1, 2, 1, 3))
f

tapply(x, f, cumsum) # list

tapply(x, f, sum) # matrix (1d)
tapply(x, f, sum, simplify = FALSE) # list

tapply(x, f) # returns f, but as integers
```

`tapply()` is just the combination of `split()` and `sapply()`. It works with data frames.

```{r, error=TRUE}
df <- data.frame(x = c(1,2,3), y = c(4,5,6), z = c(7,8,9))
f <- factor(c(1, 2, 1))
tapply(df, f, sum)
tapply(df, f, sum, simplify = FALSE)
tapply(df, f)
```

Function `by` is an object-oriented wrapper for tapply applied to data frames:

```{r, error=TRUE}
df <- data.frame(x = c(1,2,3), y = c(4,5,6), z = c(7,8,9))
f <- factor(c(1, 2, 1))
tapply(df, f, sum)
attributes(tapply(df, f, sum))

by(df, f, sum)
attributes(by(df, f, sum))
```


### `rapply()`
`rapply` is a recursive version of `lapply`. 
`rapply(object, f, classes = "ANY", deflt = NULL, how = c("unlist", "replace", "list"), ...)`

```{r, error=TRUE}
r <- list(x = list(1, 2, "a"), y = list("b", 3, "c")) 

# each numeric element that is has one added to it. Other elements remain the same
s <- rapply(r, function(x) x + 1, "numeric", how = "replace")
s

# each numeric element that is has one added to it. Other elements set to NULL
t <- rapply(r, function(x) x + 1, "numeric", how = "list")
t

# each numeric element that is has one added to it. Other elements removed
u <- rapply(r, function(x) x + 1, "numeric", how = "unlist")
u

# each numeric element that is has one added to it. Other elements set to "b"
v <- rapply(r, function(x) x + 1, "numeric", deflt = "b", how = "unlist")
v
```

### `eapply()`
`eapply` applies `FUN` to the named values from an environment and returns the results as a list. The user can request that all named objects are used (normally names that begin with a dot are not). The output is not sorted and no enclosing environments are searched.

`eapply(env, FUN, ..., all.names = FALSE, USE.NAMES = TRUE)`

```{r, error=TRUE}
env <- new.env() 
env$a <- 1:10
env$beta <- exp(-3:3)
env$logic <- c(TRUE, FALSE, FALSE, TRUE)
# what have we there?
utils::ls.str(env)

# compute the mean for each list element
eapply(env, mean)
unlist(eapply(env, mean))

# median and quartiles for each element (making use of "..." passing):
eapply(env, quantile, probs = 1:3/4)
```

### `interaction()`
`interaction` computes a factor which represents the interaction of the given factors.

```{r, error=TRUE}
interaction(c(1, 2, 2), c("A", "A", "B"))
interaction(c(1, 2, 2), c("A", "A", "B"), drop = TRUE)
```


### `ave()`
Subsets of `x[]` are averaged, where each subset consist of those observations with the same factor levels.

`ave(x, ..., FUN = mean)`

`ave()` is just `split()` and `lapply()`, and hence is essentially `tapply()` except the input vector is modified.

```{r}
ave
```

## Invoking a function
Suppose you had a list of function arguments:`args <- list(1:10, na.rm = TRUE)`. How could you then send that list to `mean()`? You need `do.call()`:

```{r}
args <- list(1:10, na.rm = TRUE)
do.call(mean, args)
# Equivalent to
mean(1:10, na.rm = TRUE)
```

## Lists of functions
A use for a list of functions is to summarise an object in multiple ways. To do that, we could store each summary function in a list, and then run them all with `lapply()`:

```{r}
x <- 1:10
funs <- list(
  sum = sum,
  mean = mean,
  median = median
)
lapply(funs, function(f) f(x))
```

### Moving lists of functions to the global environment
Three options to eliminate the use of funs$:
```{r}
funs <- list(
  sum1 = sum,
  mean1 = mean,
  median1 = median
)

funs$mean1(x)
funs$sum1(x)

with(funs, mean1(x))

attach(funs)
mean1(x)
sum1(x)
detach(funs)

list2env(funs, environment())
mean1(x)
sum1(x)
rm(list = names(funs), envir = environment())
```

## Loops that should be left as is
Some loops have no natural functional equivalent. In this section you’ll learn about three common cases:

* modifying in place
* recursive functions
* while loops

It’s possible to torture these problems to use a functional, but it’s not a good idea. You’ll create code that is harder to understand, eliminating the main reason for using functionals in the first case.

### Modifying in place
If you need to modify part of an existing data frame, it’s often better to use a for loop. For example, the following code performs a variable-by-variable transformation by matching the names of a list of functions to the names of variables in a data frame.

```{r, error=TRUE}
trans <- list(
  disp = function(x) x * 0.0163871,
  am = function(x) factor(x, labels = c("auto", "manual"))
)
for(var in names(trans)) {
  mtcars[[var]] <- trans[[var]](mtcars[[var]])
}
```

We wouldn’t normally use `lapply()` to replace this loop directly, but it is possible. Just replace the loop with `lapply()` by using `<<-`:

```{r, error=TRUE}
lapply(names(trans), function(var) {
  mtcars[[var]] <<- trans[[var]](mtcars[[var]])
})
```

The for loop is gone, but the code is longer and much harder to understand. The reader needs to understand `<<`- and how `x[[y]] <<- z works` (it’s not simple!).

### Recursive relationships
It’s hard to convert a for loop into a functional when the relationship between elements is not independent, or is defined recursively. For example, exponential smoothing works by taking a weighted average of the current and previous data points. 

### While loops
Another type of looping construct in R is the `while` loop. It keeps running until some condition is met. `while` loops are more general than for loops: you can rewrite every `for` loop as a `while` loop, but you can’t do the reverse. For example, we could turn this for loop:

```{r}
for (i in 1:10) print(i)

i <- 1
while(i <= 10) {
  print(i)
  i <- i + 1
}
```

Not every `while` loop can be turned into a `for` loop because many `while` loops don’t know in advance how many times they will be run:

```{r}
i <- 0
while(TRUE) {
  if (runif(1) > 0.9) break
  i <- i + 1
}
```

This is a common problem when you’re writing simulations.




