---
title: "Functions"
output:
  html_document:
    number_sections: yes
    theme: sandstone
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(tidy=TRUE, fig.align='center')
```

```{r, echo=FALSE, include=FALSE}
library(lobstr)
library(pryr)
library(knitr)
library(kableExtra)
library(magrittr)
library(dplyr)
library(tidyverse)
```

# Functions
The most important thing to understand about R is that functions are objects in their own right, a language property often called “first-class functions”. You can work with them exactly the same way you work with any other type of object. 

## Function components
All R functions have three parts:

* The `formals()`, the list of arguments that control how you call the function.

* The `body()`, the code inside the function.

* The `environment()`, which determines how the function finds values associated with names.

When you print a function in R, it shows you these three important components. If the environment isn’t displayed, it means that the function was created in the global environment.

```{r}
f <- function(x) {
  # A comment
  x ^ 2
}
f
formals(f)
body(f)
environment(f)
```

Like all objects in R, functions can also possess any number of additional `attributes()`. One attribute used by base R is “srcref”, short for source reference, which points to the source code used to create the function. Unlike `body()`, this contains code comments and other formatting. You can also add attributes to a function. For example, you can set the `class()` and add a custom `print()` method.

```{r}
attributes(f)
```

The assignment forms of `body()`, `formals()`, and `environment()` can also be used to modify functions.

```{r}
formals(f) <- alist(x = , y = 5 * 2) # alist allows tags without values. Values are not evaluated
f
body(f) <- expression(x ^ 3 + y)
f
f(2)
```

## Primitive functions
There is one exception to the rule that functions have three components. Primitive functions, like `sum()`, call C code directly with `.Primitive()` and contain no R code. Therefore their `formals()`, `body()`, and `environment()` are all `NULL`:

```{r}
sum
formals(sum)
body(sum)
environment(sum)
```

Primitive functions are only found in the base package. The chief advantage of primitive functions is that they can truly modify objects in place

## Exercises
1. What function allows you to tell if an object is a function? What function allows you to tell if a function is a primitive function?  
**A**: 
    ```{r, results='hold'}  
    f <- function(x) x ^ 2
    is.function(f)
    is.function(sum)
    is.primitive(f)
    is.primitive(sum)
    ```

2. This code makes a list of all functions in the base package. Use it to answer the following questions:
    ```{r}
    objs <- mget(ls("package:base"), inherits = TRUE)
    funs <- Filter(is.function, objs)
    ```

    Which base function has the most arguments?
    ```{r}
    f_arg_length <- funs %>% map_int(~ length(formals(.)))
    f_arg_length[which.max(f_arg_length)]
    ```

    How many base functions have no arguments? What’s special about those functions?
    ```{r}
    # functions without arguments
    f_arg_length <- funs %>% map_int(~ length(formals(.)))
    f_arg_length[which.max(f_arg_length)]
    sum(f_arg_length == 0) 
    
    # all primitive functions return NULL for formals, but so do other functions with no arguments
    f_arg_NULL <- funs %>% map_lgl(~ is.null(formals(.)))
    sum(f_arg_NULL) 
    
    f_arg_0 <- funs %>% map_lgl(~ !is.null(formals(.)) && length(formals(.)) == 0)
    sum(f_arg_0) #0 - non-primitive functions with no arguments return NULL for a formal
    
    f_arg_primitve <- funs %>% map_lgl(~ is.primitive(.))
    sum(f_arg_primitve) # number of primitive functions
    
    f_arg_NULL_not_primitive <- funs %>% map_lgl(~ is.null(formals(.)) && !is.primitive(.))
    sum(f_arg_NULL_not_primitive) # number of non-primitive functions with no arguments
    
    # non-primitve functions with no arguments
    str(funs[which(f_arg_NULL_not_primitive)])
    ```

    How could you adapt the code to find all primitive functions?
    ```{r}
    objs <- mget(ls("package:base"), inherits = TRUE)
    funs <- Filter(is.primitive, objs)
    length(funs)
    ```

3. What are the three important components of a function?  
**A**: The `formals()`, `body()` and `environment()`. 

4. When does printing a function not show what environment it was created in?  
**A**: When the function was created in the global environment.


## Lexical scoping




# DO RECURSIVE SUBSETTING
