---
title: "Quasiquotation"
output:
  html_document:
    number_sections: yes
    theme: sandstone
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(tidy=TRUE, fig.align='center')
```

```{r, echo=FALSE, include=FALSE}
library(rlang)
library(pryr)
library(knitr)
library(kableExtra)
library(magrittr)
library(dplyr)
library(tidyverse)
library(lobstr)
library(memoise)
library(purrr)
```

# Quasiquotation
There are two sides to quasiquotation:

* Quotation allows you to capture the AST associated with an argument. As a function author, this gives you a lot of power to influence how expressions are evaluated.

* Unquotation allows you to selectively evaluate parts of a quoted expression. This is a powerful tool that makes it easy to build up a complex AST from simpler fragments.

## Example
Imagine you’re creating a lot of strings by joining together words:

```{r}
paste("Good", "morning", "Hadley")
paste("Good", "afternoon", "Alice")
```

You are sick and tired of writing all those quotes, and instead you just want to use bare words. To that end, you’ve managed to write the following function:

```{r}
cement <- function(...) {
  dots <- exprs(...)
  paste(purrr::map(dots, expr_name), collapse = " ")
}

cement(Good, morning, Hadley)
cement(Good, afternoon, Alice)
```

Formally, this function quotes the arguments in `...`. You can think of it as automatically putting quotation marks around each argument. That’s not precisely true as the intermediate objects it generates are expressions, not strings, but it’s a useful approximation for now.

This function is nice because we no longer need to type quotes. The problem, however, comes when we want to use variables. It’s easy to use variables with `paste()` as we just don’t surround them with quotes:

```{r}
name <- "Hadley"
time <- "morning"

paste("Good", time, name)
```

Obviously this doesn’t work with `cement()` because every input is automatically quoted:

```{r}
cement(Good, time, name)
```

We need some way to explicitly unquote the input, to tell `cement()` to remove the automatic quote marks. Here we need time and name to be treated differently to Good. Quasiquotation give us a standard tool to do so: `!!`, called “unquote”, and pronounced bang-bang. `!!` tells a quoting function to drop the implicit quotes:

```{r}
cement(Good, !!time, !!name)
```

It’s useful to compare `cement()` and `paste()` directly. `paste()` evaluates its arguments, so we need to quote where needed; `cement()` quotes its arguments, so we need to unquote where needed.

```{r}
paste("Good", time, name)
cement(Good, !!time, !!name)
```

### Other
```{r}
name <- "tom"
time <- "morning"

dots <- exprs(Good, morning, tom)
typeof(dots[[2]])
paste(purrr::map(dots, expr_name), collapse = " ")

dots <- exprs(Good, time, name)
typeof(dots[[2]])
paste(purrr::map(dots, expr_name), collapse = " ")

dots <- exprs(Good, !!time, !!name)
typeof(dots[[2]])
paste(purrr::map(dots, expr_name), collapse = " ")

time <- 33
dots <- exprs(Good, !!time, !!name)
typeof(dots[[2]])
paste(purrr::map(dots, expr_name), collapse = " ")
```

## Vocabulary
The distinction between quoted and evaluated arguments is important:

* An evaluated argument obeys R’s usual evaluation rules.
* A quoted argument is captured by the function and something unusual will happen.

If you’re even unsure about whether an argument is quoted or evaluated, try executing the code outside of the function. If it doesn’t work, then that argument is quoted. For example, you can use this technique to determine that the first argument to `library()` is quoted:

```{r, eval=FALSE}
# works
library(MASS)

# fails
MASS
```

Talking about whether an argument is quoted or evaluated is a more precise way of stating whether or not a function uses NSE. 

### Exercises
1. For each function in the following base R code, identify which arguments are quoted and which are evaluated.  
    ```{r, eval=FALSE}
    library(MASS)

    mtcars2 <- subset(mtcars, cyl == 4)
    
    with(mtcars2, sum(vs))
    sum(mtcars2$am)
    
    rm(mtcars2)
    ```
**A**:
    ```{r, error=TRUE}
    library(MASS)
    MASS # error - quoted
    
    mtcars2 <- subset(mtcars, cyl == 4)
    mtcars
    cyl # error - quoted
    
    with(mtcars2, sum(vs))
    vs # error - quoted
    
    sum(mtcars2$am)
    mtcars2$am
    ```

2. For each function in the following tidyverse code, identify which arguments are quoted and which are evaluated.
    ```{r, eval = FALSE}
    library(dplyr)
    library(ggplot2)
    
    by_cyl <- mtcars %>%
      group_by(cyl) %>%
      summarise(mean = mean(mpg))
    
    ggplot(by_cyl, aes(cyl, mean)) + geom_point()
    ```
**A**:
    ```{r, eval=FALSE}
    by_cyl <- mtcars %>%
      group_by(cyl) %>%  # cyl - quoted
      summarise(mean = mean(mpg)) # mean = mean(mpg) - quoted
    
    ggplot(by_cyl, aes(cyl, mean)) + geom_point() # cyl, mean - quoted
    ```

## Quotation
The first part of quasiquotation is quotation: capturing an AST without evaluating it. There are two components to this: capturing an expression directly, and capturing an expression from a lazily-evaluated function argument.

### With rlang
For interactive exploration, the most important quoting function is `expr()`. It captures its argument exactly as provided:

```{r}
expr(x + y)
expr(1 / 2 / 3)

expr(x+y)[[1]]
expr(x+y)[[2]]
expr(x+y)[[3]]
```

`exprs()` is useful interactively to make a list of expressions:

```{r}
exprs(x = x ^ 2, y = y ^ 3, z = z ^ 4)
```

`expr()` is great for interactive exploration, because it captures what you, the developer, typed. It’s not useful inside a function:

```{r}
f1 <- function(x) expr(x)
f1(a + b + c)
```

Instead, we need another function: `enexpr()`. This captures what the user supplies to the function by looking at the internal promise object that powers lazy evaluation.

```{r}
f2 <- function(x) enexpr(x)
f2(a + b + c)
```

To capture multiple arguments, use `enexprs()`:

```{r}
f <- function(...) enexprs(...)
f(x = 1, y = 10 * z)
```

Note: `exprs()` can return missing arguments, but no`expr()`
```{r, error=TRUE}
val <- exprs(x = )
is_missing(val$x)

val <- expr(x = )
```

### With Base R
The base equivalent of `expr()` is `quote()`:

```{r}
quote(x + y)
quote(1 / 2 / 3)

quote(x + y)[[1]]
```

It is identical to `expr()` except that does not support unquoting, so it is a quoting function, not a quasiquoting function.

The base function closest to `enexpr()` is `substitute()`:

```{r}
f <- function(x) substitute(x)
f(x + y + z)
```

You’ll most often see it used to capture unevaluated arguments; often in concert with `deparse()` to create labels for output. However, `substitute()` also does “substitution”: if you give it an expression, rather than a symbol, it will substitute in values of symbols defined in the current environment.

```{r}
f <- function(x) substitute(x * 2)
f(a + b + c)
```

But:

```{r, error=TRUE}
f <- function(x) enexpr(x * 2)
f(a + b + c)
```

`substitute()` provides a sort of automatic unquoting for any symbol that is bound to a value. However, making use of this behaviour can make for hard to read code, because for example, taken out of context, you can’t tell if the goal of `substitute(x + y)` is to replace `x`, or, `y`, or both. If you do want to use `substitute()` in this way, I recommend that you use the 2nd argument to make it clear that is your goal:

```{r}
substitute(x * y * z, list(x = 10, y = quote(a + b)))
```

The base equivalent to `exprs()` is `alist()`:

```{r}
alist(x = 1, y = x + 2)
exprs(x = 1, y = x + 2)
```

Finally, there is no function in base R that is equivalent to `enexprs()` but you can reproduce its behaviour with `eval(substitute(alist(...)))`. 

```{r}
f <- function(...) enexprs(...)
f(x = 1, y = 10 * z)

f1 <- function(...) eval(substitute(alist(...)))
f1(x = 1, y = 10 * z)
```

There are two other important base quoting functions that we’ll cover elsewhere:

* `bquote()` provides a limited form of quasiquotation, and is discussed in unquoting with base R.
* `~`, the formula, is a quoting function that also captures the environment. It’s the inspiration for quosures, the topic of the next chapter.

### Exercises
1. What happens if you try to use `enexpr()` with an expression? What happens if you try to use `enexpr()` with a missing argument?  
**A**:
    ```{r}
    f <- function(x) enexpr(x)
    
    f(expr(x)) # return expr(x), a call
    class(f(expr(x)))
    
    f() # returns the missing arg
    is_missing(f())
    ```

2. Compare and contrast the following two functions. Can you predict the output before running them?  
    ```{r}
    f1 <- function(x, y) {
      exprs(x = x, y = y)
    }
    f2 <- function(x, y) {
      enexprs(x = x, y = y)
    }
    f1(a + b, c + d)
    f2(a + b, c + d)
    ```
**A**: Both return a named list of expressions. `f1()` will return the arguments supplied to `exprs()` within the body of `f1()`. `f2()` will return the arguments supplied to `f2()`.

3. How are `exprs(a)` and `exprs(a = )` different? Think about both the input and the output.  
**A**: `exprs(a)` returns an unamed list of one element, the symbol `a`. `exprs(a = )` return a list of one element, the missing argument, named `a`.
    ```{r}
    exprs(a)
    exprs(a = )
    ```

4. What does the following command return? What information is lost? Why?
    ```{r}
    expr({
      x +              y # comment  
    })
    ```
**A**: Whitespace and comments are not retained. However, you can get the source information from the attributes of the expresssion.  
    ```{r}
    attributes(expr({
      x +              y # comment
    }))
    ```

5. The documentation for `substitute()` says:

    *Substitution takes place by examining each component of the parse tree as follows: If it is not a bound symbol in env, it is unchanged. If     it   is a promise object, i.e., a formal argument to a function or explicitly created using* `delayedAssign()`*, the expression slot of the     promise replaces the symbol. If it is an ordinary variable, its value is substituted, unless env is* `.GlobalEnv` *in which case the            symbol is left unchanged.*

    Create four examples that illustrate each of the different cases.
**A**:
    If it is not a bound symbol in env, it is unchanged
    ```{r}
    substitute(x, list(y = 1))
    ```
    
    If it is a promise object as a formal argument to a function, the expression slot of the promise replaces the symbol.
    ```{r}
    f <- function(x) substitute(x)
    f(y) # y 
    
    # BUT NOTE:
    f <- function(x) {x <- 1; substitute(x)}
    f(y) # 1 - env defaults to current evaluation environment
    
    f <- function(x) {substitute(x, list(NULL))}
    f(y) # x - env defaults evaluation environment with no bound symbols
    ```
    
    If it is a promise object explicitly created using `delayedAssign()`, the expression slot of the promise replaces the symbol.
    ```{r}
    y <- 1
    delayedAssign("x", y)
    substitute(x) # shows only 'x', as it is in the global env.
    x # 1
    rm(x)
    
    e <- new.env()
    e$z <- 1
    delayedAssign("x", z, eval.env = e, assign.env = e)
    e$x # 1
    substitute(x, e) # shows z
    eval(substitute(x, e), e) # 1
    ```
    
    If it is an ordinary variable, its value is substituted, unless env is `.GlobalEnv` in which case the symbol is left unchanged.
    ```{r}
    substitute(x, list(x = 1)) # x is substituted
    
    e <- new.env()
    e$x <- 1
    substitute(x, e) # x is substituted
    
    x <- 1
    substitute(x, .GlobalEnv) # x is not substituted
    ```

## Evaluation
Typically you have quoted a function argument for one of two reasons:

* You want to operate on the AST using the techniques described in the previous chapter.
* You want to run, or evaluate the code in a special context, as described in depth next chapter.

Evaluation is a rich topic, so we’ll cover it in depth in the next chapter. Here I’ll just illustrate the most important ideas. The most important base R function is `base::eval()`. Its first argument is the expression to evalute:

```{r}
ru5 <- expr(runif(5))
ru5

eval(ru5)
eval(ru5)
```

Note that every time we evaluate this expression we get a different result.

The second argument to `eval()` is the environment in which the expression is evaluated. Manipulating this environment gives us amazing power to control the execution of R code. 

```{r}
x <- 9
fx <- expr(f(x))

eval(fx, env(f = function(x) x * 10))
eval(fx, env(f = function(x) x ^ 2))

eval(fx, env(f = function(x) {print(env_parent()); x * 10})) 
# parent of function evaluation is Global Env, hence can find x
# env() creates a child of the current environment by default and takes a variable number of named objects to populate it.
```

## Unquotation
Evaluation is a developer tool: in combination with quoting, it allows the author of a function to capture an argument and evaluate it in a special way. Unquoting is related to evaluation, but it’s a user tool: it allows the person calling the function to selectively evaluate parts of the expression that would otherwise be quoted.

### With rlang
All quoting functions in rlang (`expr()`, `enexpr()`, and friends) support unquoting with `!!` (called “unquote”, and pronounced bang-bang) and `!!!` (called “unquote-splice”, and pronounced bang-bang-bang). They both replace nodes in the AST. `!!` is a one-to-one replacement. It takes a single expression and inlines the AST at the location of the `!!`.

```{r}
x <- expr(a + b + c)
expr(f(!!x, y))
expr(f(x, y))

lobstr::ast(f(!!x, y))
lobstr::ast(f(x, y))
```

`!!!` is a one-to-many replacement. It takes a list of expressions and inserts them at the location of the `!!!`:

```{r}
x <- exprs(1, 2, 3, y = 10)
expr(f(!!!x, z = z))
lobstr::ast(f(!!!x, z = z))
```

### The polite fiction of `!!`
So far we have acted as if `!!` and `!!!` are regular prefix operators like `+` , `-`, and `!`. They’re not. Instead, from R’s perspective, `!!` and `!!!` are simply the repeated application of `!`:

```{r}
!!TRUE # TRUE
!!!TRUE # FALSE
```

`!!` and `!!!` have special behaviour inside all quoting functions powered by rlang, and the unquoting operators are given precedence similar to `+` and `-`, not `!`. We do this because the operator precedence for `!` is surprisingly low: it has lower precedence than that of the binary algebraic and logical operators. Most of the time this doesn’t matter as it is unusual to mix `!` and binary operators (e.g. you typically would not write `!x + y` or `!x > y`). However, expressions like `!!x + !!y` are not uncommon when unquoting, and requiring explicit parentheses, `(!!x) + (!!y)`, feels onerous. For this reason, rlang manipulates the AST to give the unquoting operators a higher, more natural, precedence. (And if for some reason you do need to doubly negate a value in a quasiquoting function, you can just add parentheses `!(!x)`.)

The biggest downside to using a fake operator is that you might get silent errors when misusing `!!` outside of quasiquoting functions. Most of the time this is not an issue because `!!` is typically used to unquote expressions or quosures. Since expressions are not supported by the negation operator, you will get an argument type error in this case:

```{r, error=TRUE}
x <- quote(variable)
!!x

y <- expr(variable)
!!y
```

However be extra careful when unquoting numeric values that can be negated silently:

```{r}
x <- 100
with(mtcars, cyl + !!x)
with(mtcars, cyl)
```

Instead of adding the value of `x` to `cyl` as intended, we have in fact added the double negation of `x`:

```{r}
!x
!!x
```

### With Base R
Base R has one function that implements quasiquotation: `bquote()`. It uses `.()` for unquoting:

```{r}
xyz <- bquote((x + y + z))
bquote(-.(xyz) / 2)

# Also
substitute(-XYZ / 2, list(XYZ = xyz))
```

`bquote()` is a neat function, but is not used by any other function in base R. Instead functions that quote an argument use some other technique to allow indirect specification. There are four basic forms seen in base R:

1. A pair of quoting and non-quoting functions. For example, `$` has two arguments, and the second argument is quoted. This is easier to see if you write in prefix form: `mtcars$cyl` is equivalent to `$(mtcars, cyl)`. If you want to refer to a variable indirectly, you use `[[`, as it takes the name of a variable as a string.
    ```{r}
    x <- list(var = 1, y = 2)
    var <- "y"
    
    x$var # 1
    x[[var]] # 2
    ```
`<-/assign()` work similarly:
    ```{r}
    y <- 1
    x <- y
    x

    var <- "x"
    assign(var, 3)
    x

    # note
    z <- 3
    "x" <- z
    x
    ```

2. A pair of quoting and non-quoting arguments. For example, `data()`, `rm()`, and `save()` allow you to provide bare variable names in `...`, or a character vector of variable names in `list`:
    ```{r}
    x <- 1
    rm(x)
    
    y <- 2
    vars <- c("y", "vars")
    rm(list = vars)
    ```

3. An argument that controls whether a different argument is quoting or non-quoting. For example, in `library()`, the character.only argument controls the quoting behaviour of of the first argument, package:
    ```{r, eval=FALSE}
    library(MASS)
    
    pkg <- "MASS"
    library(pkg, character.only = TRUE)
    ```
`demo()`, `detach()`, `example()`, and `require()` work similarly.

4. Quoting if evaluation fails. For example, the first argument to `help()` is non-quoting if it evaluates to a string; if evaluation fails, the first argument is quoted.
    ```{r, eval=FALSE}
    # Shows help for var
    help(var)
    
    var <- "mean"
    # Shows help for mean
    help(var)
    
    var <- 10
    # Shows help for var
    help(var)
    ```
`ls()`, `page()`, and `match.fun()` work similarly.

Some quoting functions, like `subset()`, `transform()`, and `with()`, don’t have a non-quoting form. This is because they are seen as wrappers around `[` and `[<-` that are only suitable for interactive use.

### Missing arguments
Occasionally it is useful to unquote a missing argument, but the naive approach doesn’t work:

```{r, error=TRUE}
arg <- missing_arg()
expr(foo(!!arg, !!arg))
```

You can either wrap in a list and use unquote-splice, or use the `maybe_missing()` helper:

```{r}
args <- list(missing_arg(), missing_arg())
expr(foo(!!!args))
expr(foo(!!maybe_missing(arg), !!maybe_missing(arg)))
```

### Exercises
1. Given the following components:
    ```{r}
    xy <- expr(x + y)
    xz <- expr(x + z)
    yz <- expr(y + z)
    abc <- exprs(a, b, c)
    ```
Use quasiquotation to construct the following calls:
    ```{r, eval=FALSE}
    (x + y) / (y + z)
    -(x + z) ^ (y + z)
    (x + y) + (y + z) - (x + y)
    atan2(x + y, y + z)
    sum(x + y, x + y, y + z)
    sum(a, b, c)
    mean(c(a, b, c), na.rm = TRUE)
    foo(a = x + y, b = y + z)
    ```
**A**: 
    ```{r}
    expr(!!xy / !!yz)
    expr(-(!!xz) ^ !!yz)
    expr(!!xy + !!yz - !!xy)
    expr(atan2(!!xy, !!yz))
    expr(sum(!!xy, !!xy, !!yz))
    expr(sum(!!!abc))
    expr(sum(!!!abc, na.rm = TRUE))
    expr(foo(a = !!xy, b = !!yz))
    ```

2. Explain why both `!0 + !0` and `!1 + !1` return `FALSE` while `!0 + !1` returns TRUE.  
**A**: `!0 + !0 = !(0 + !0) = 0`. `!1 + !1 = !(1 + !1) = 0`. `!0 + !1 = !(0 + !1) = 1`

3. Base functions `match.fun()`, `page()`, and `ls()` all try to automatically determine whether you want standard or non-standard evaluation. Each uses a different approach. Figure out the essence of each approach by reading the source code, then compare and contrast the techniques.  
**A**: 
* `match.fun()` - 


4. The following two calls print the same, but are actually different:
    ```{r}
    (a <- expr(mean(1:10)))
    (b <- expr(mean(!!(1:10))))
    
    identical(a, b)
    ```
What’s the difference? Which one is more natural?  
**A**: in the 2nd expression `!!(1:10)` evaluates to a sequence of integers, but this prints as 1:10. In the first expression, `1:10` is a call, `:`(1, 10)`.
    ```{r}
    (a <- expr(mean(1:10)))
    (b <- expr(mean(!!(1:10))))
    eval(a)
    eval(b)
    
    a[[1]]
    b[[1]]
    
    typeof(a[[1]]) # symbol
    typeof(b[[1]]) # symbol
    
    a[[2]]
    b[[2]]
    
    typeof(a[[2]]) # language
    typeof(b[[2]]) # integer
    
    class(a[[2]]) # call
    class(b[[2]]) # integer
    
    
    lobstr::ast(!!a)
    lobstr::ast(!!b)  
    ```




