---
title: "S3"
output:
  html_document:
    number_sections: yes
    theme: sandstone
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(tidy=TRUE, fig.align='center')
```

```{r, echo=FALSE, include=FALSE}
library(rlang)
library(pryr)
library(knitr)
library(kableExtra)
library(magrittr)
library(dplyr)
library(tidyverse)
library(lobstr)
library(memoise)
library(purrr)
library(sloop)
```

# Introduction
S3 is R’s first and simplest OO system. S3 is informal and ad hoc, but it has a certain elegance in its minimalism: you can’t take away any part of it and still have a useful OO system. Because of these reasons, S3 should be your default choice for OO programming: you should use it unless you have a compelling reason otherwise. S3 is the only OO system used in the base and stats packages, and it’s the most commonly used system in CRAN packages.

S3 is a very flexible system: it allows you to do a lot of things that are quite ill-advised. If you’re coming from a strict environment like Java, this will seem pretty frightening (and it is!) but it does give R programmers a tremendous amount of freedom. While it’s very difficult to prevent someone from doing something you don’t want them to do, your users will never be held back because there is something you haven’t implemented yet. Since S3 has few built-in constraints, the key to its successful use is applying the constraints yourself. This chapter will teach you the conventions you should (almost) always adhere to in order to use S3 safely.

# Basics
An S3 object is built on top of a base type with the “class” attribute set. The base type is typically a vector, although we will see later that it’s possible to use other types of classes. For example, take the factor. It is built on top of an integer vector, and the value of the class attribute is “factor”. It stores information about the “levels” in another attribute.

```{r}
f <- factor("a")

typeof(f)
attributes(f)
```

An S3 object behaves differently from its underlying base type because of **generic functions**, or **generics** for short. A generic executes different code depending on the class of one of its arguments, almost always the first. You can see this difference with the most important generic function: `print()`.

```{r}
print(f)
print(unclass(f))
```

`unclass()` strips the class attribute from its input, so it is a useful tool for seeing what special behaviour an S3 class adds.

`str()` shows the internal structure of S3 objects. Be careful when using `str()`: some S3 classes provide a custom `str()` method which can hide the underlying details. For example, take the `POSIXlt` class, which is one of the two classes used to represent date-time data:

```{r}
time <- strptime("2017-01-01", "%Y-%m-%d")
str(time)
str(unclass(time))
```

A **generic** and its **methods** are functions that operate on classes. The role of a generic is to find the right method for the arguments that it is provided, the process of **method dispatch**. A method is a function that implements the generic behaviour for a specific class. In other words the job of the generic is to find the right method; the job of the method is to do the work.

S3 methods are functions with a special naming scheme, `generic.class()`. For example, the Date method for the `mean()` generic is called `mean.Date()`, and the factor method for `print()` is called `print.factor()`. This is the reason that most modern style guides discourage the use of `.` in function names: it makes them look like S3 methods. For example, is `t.test()` the t method for `test` objects?

You can find some S3 methods (those in the base package and those that you’ve created) by typing their names. However, this will not work with most packages because S3 methods are not exported: they live only inside the package, and are not available from the global environment. Instead, you can use `getS3method()`, which will work regardless of where the method lives:

```{r}
# Only works because the method is in the base package
mean.Date

# Always works
getS3method("mean", "Date")
```

## Exercises
1. The most important S3 objects in base R are factors, data frames, and date/times (Dates, POSIXct, POSIXlt). You’ve already seen the attributes and base type that factors are built on. What base types and attributes are the others built on?  
**A**: Data frames are build up on (named) lists. Together with the `row.names` attribute and after setting the class to “data.frame”, we get a classical data frame.
    ```{r}
    df <- data.frame()
    attributes(df)
    
    df_build <- structure(list(1:2, 3:4),
                          names = c("a", "b"),
                          row.names = 1:2, 
                          class = "data.frame")
    
    df_classic <- data.frame(a = 1:2, b = 3:4)
    
    identical(df_build, df_classic)
    ```
Date is just a double with the class attribute set to “Date”

    ```{r}
    d <- as.Date("1970-01-01")
    attributes(d)
    
    date_build <- structure(0, class = "Date")
    date_classic <- as.Date("1970-01-01")
    identical(date_build, date_classic)
    ```
POSIXct is a class for date/times that inherits from POSIXt and is built on doubles as well. The only attribute is tz (for timezone)
    ```{r}
    POSIXct_build <- structure(1, class = c("POSIXct", "POSIXt"), tzone = "CET")
    POSIXct_classic <- .POSIXct(1, tz = "CET") # note that tz's default is NULL
    identical(POSIXct_build, POSIXct_classic)
    ```
POSIXlt is another date/time class that inherits from POSIXt. It is built on top of a named list and a tzone attribute. Differences between POSIXct and POSIXlt are described in `?DateTimeClasses`.
    ```{r}
    POSIXlt_build <- structure(list(sec = 30,
                                    min = 30L,
                                    hour = 14L,
                                    mday = 1L,
                                    mon = 0L,
                                    year = 70L,
                                    wday = 4L,
                                    yday = 0L,
                                    isdst = 0L,
                                    zone = "CET",
                                    gmtoff = 3600L),
                               tzone = c("", "CET", "CEST"),
                               class = c("POSIXlt", "POSIXt"))
    POSIXlt_classic <- as.POSIXlt(.POSIXct(13.5 * 3600 + 30))
    identical(POSIXlt_build, POSIXlt_classic)
    ```

2. Describe the difference in behaviour in these two calls.
    ```{r}
    set.seed(1014)
    some_days <- as.Date("2017-01-31") + sample(10, 5)
    
    mean(some_days)
    mean(unclass(some_days))
    ```
**A**: Since `mean()` is a generic and some_days is an object of class Date, the first call results in `mean.Date(some_days)`. In the second call `unclass()` removes the class attribute from some_days, which means that unclass(some_days)` is not an OO object and the call results in `mean.default(unclass(some_days))`, which calculates the mean of the underlying double. Note:
    ```{r}
    mean.Date
    ```

# Classes
S3 is a simple and ad hoc system, and has no formal definition of a class. To make an object an instance of a class, you simply take an existing object and set the class attribute. You can do that during creation with `structure()`, or after the fact with `class<-()`:

```{r}
# Create and assign class in one step
foo <- structure(list(), class = "foo")

# Create, then set class
foo <- list()
class(foo) <- "foo"
```

You can determine the class of any object using `class(x)`, and see if an object inherits from a specific class using `inherits(x, "classname")`.

```{r}
class(foo)
inherits(foo, "foo")
```

It’s possible to provide a vector of class names, which allows S3 to implement a basic style of inheritance. This allows you to reduce your workload by allowing classes to share code where possible. We’ll come back to this idea in inheritance.

S3 has no checks for correctness. This means you can change the class of existing objects:

```{r}
# Create a linear model
mod <- lm(log(mpg) ~ log(disp), data = mtcars)
class(mod)
print(mod)

# Turn it into a data frame (?!)
class(mod) <- "data.frame"
# Unsurprisingly this doesn't work very well
print(mod)
```

When creating your own class, you should always provide:

* A **constructor**, `new_x()`, that efficiently creates new objects with the correct structure.

For more complicated classes, you may also want to provide:

* A **validator**, `validate_x()`, that performs more expensive checks that the object has correct values.
* A **helper**, `x()`, that provides a convenient and neatly parameterised way for others to construct and validate (create) objects of this class.

## Constructors
S3 doesn’t provide a formal definition of a class, so it has no built-in way to ensure that all objects of a given class have the same structure (i.e. same attributes with the same types). Instead, you should enforce a consistent structure yourself by using a constructor. A constructor is a function whose job is to create objects of a given class, ensuring that they always have the same structure.

There are three rules that a constructor should follow. It should:

* Be called `new_class_name()`.
* Have one argument for the base object, and one for each attribute. (More if the class can be subclassed, see inheritance.)
* Check the types of the base object and each attribute.

Base R generally does not provide constructors (three exceptions are the internal `.difftime()`, `.POSIXct()`, and `.POSIXlt()`) so we’ll demonstrate constructors by filling in some missing pieces in base. (If you want to use these constructors in your own code, you can use the versions exported by the sloop package, which complete a few details that we skip here in order to focus on the core issues.)

We’ll start with one of the simplest S3 classes in base R: Date, which is just a double with a class attribute. The constructor rules lead to the slightly awkward name `new_Date()`, because the existing base class uses a capital letter. I recommend using lower case class names to avoid this problem.

```{r}
new_Date <- function(x) {
  stopifnot(is.double(x))
  structure(x, class = "Date")
}

new_Date(c(-1, 0, 1))
```

A slightly more complicated example is POSIXct, which is used to represent date-times. It is again built on a double, but has an attribute that specifies the time zone, a length 1 character vector. R defaults to using the local time zone, which is represented by the empty string. To create the constructor, we need to make sure each attribute of the class gets an argument to the constructor. This gives us:

```{r}
new_POSIXct <- function(x, tzone = "") {
  stopifnot(is.double(x))
  stopifnot(is.character(tzone), length(tzone) == 1)
  
  structure(x, 
    class = c("POSIXct", "POSIXt"),
    tzone = tzone
  )
}

new_POSIXct(1)
new_POSIXct(1, tzone = "UTC")
```

Note that POSIXct uses a class *vector*; we’ll come back to what that means in inheritance.

Generally, the constructor should not check that the values are valid because such checks are often expensive.

## Validators
More complicated classes will require more complicated checks for validity. Take factors, for example. The constructor function only checks that the structure is correct:

```{r}
new_factor <- function(x, levels) {
  stopifnot(is.integer(x))
  stopifnot(is.character(levels))
  
  structure(
    x,
    levels = levels,
    class = "factor"
  )
}
```

So it’s possible to use this to create invalid factors:

```{r, error=TRUE}
new_factor(1:5, "a")
new_factor(0:1, "a")
```

Rather than encumbering the constructor with complicated checks, it’s better to put them in a separate function. This is a good idea because it allows you to cheaply create new objects when you know that the values are correct, and to re-use the checks in other places.

```{r, error=TRUE}
validate_factor <- function(x) {
  values <- unclass(x)
  levels <- attr(x, "levels")
  
  if (!all(!is.na(values) & values > 0)) {
    stop(
      "All `x` values must be non-missing and greater than zero",
      call. = FALSE
    )
  }
  
  if (length(levels) < max(values)) {
    stop(
      "There must at least as many `levels` as possible values in `x`",
      call. = FALSE
    )
  }
  
  x
}

validate_factor(new_factor(1:5, "a"))
validate_factor(new_factor(0:1, "a"))
```

## Helpers
If you want others to construct objects from your class, you should also provide a helper method that makes their life as easy as possible. This should have the same name as the class, and should be parameterised in a convenient way. `factor()` is a good example of this as well: you want to automatically derive the internal representation from a vector. The simplest possible implementation looks something like this:

```{r, error=TRUE}
factor <- function(x, levels = unique(x)) {
  ind <- match(x, levels)
  validate_factor(new_factor(ind, levels))
}
factor(c("a", "a", "b"))
factor(c("a", "a", "b"), levels = "a")
```

## Object styles
S3 gives you the freedom to build a new class on top of any existing base type. So far, we’ve focussed on vector-style where you take an existing vector type and add some attributes. Importantly, a single vector-style object represents multiple values. There are two other important styles: scalar-style and data-frame-style.

Each **scalar**-style object represents a single “value”, and are built on top of named lists. This is the style that you are most likely to use in practice. The constructor for the scalar type is slightly different because the arguments become named elements of the list, rather than attributes.

```{r}
new_scalar_class <- function(x, y, z) {
  structure(
    list(
      x = x,
      y = y,
      z = z
    ),
    class = "scalar_class"
  )
}
```

In base R, the most important example of this style is `lm`, the class returned when you fit a linear model:

```{r}
mod <- lm(mpg ~ wt, data = mtcars)
typeof(mod)
names(mod)
class(mod)
attributes(mod)
```

The **data-frame**-style builds on top of a data frame (a named list where each element is a vector of the same length), and adds additional attributes to store important metadata. A data-frame-style constructor looks like:

```{r}
new_df_class <- function(df, attr1, attr2) {
  stopifnot(is.data.frame(df))
  
  structure(
    df, 
    attr1 = attr1,
    attr2 = attr2,
    class = c("df_class", "data.frame")
  )
}
```

Collectively, we’ll call the attributes of a vector-style or data-frame-style class and the names of a list-style class the **fields** of an object.

## Exercises


# Generics and methods
The job of an S3 generic is to perform method dispatch, i.e. find the function designed to work specifically for the given class. S3 generics have a simple structure: they call `UseMethod()`, which then calls the right method. `UseMethod()` takes two arguments: the name of the generic function (required), and the argument to use for method dispatch (optional). If you omit the second argument it will dispatch based on the first argument, which is what I generally advise.

```{r}
# Dispatches on x
generic <- function(x, y, ...) {
  UseMethod("generic")
}

# Dispatches on y
generic2 <- function(x, y, ...) {
  UseMethod("generic2", y)
}
```

Note that you don’t pass any of the arguments of the generic to `UseMethod()`; it uses black magic to pass them on automatically. Generally, you should avoid doing any computation in a generic, because the semantics are complicated and few people know the details. In general, any modifications to the arguments of the generic will be undone, leading to much confusion.

A generic isn’t useful without some methods, which are just functions that follow a naming scheme (`generic.class`). Because a method is just a function with a special name, you can call methods directly, but you generally shouldn’t. The main reason to call the method directly is that it sometimes leads to considerable performance improvements. See performance for an example.

```{r}
generic.foo <- function(x, y, ...) {
  message("foo method")
}

str(new_s3_scalar(class = "foo"))
generic(new_s3_scalar(class = "foo"))
```

You can see all the methods defined for a generic with `s3_methods_generic()` or `utils::methods()`:

```{r}
s3_methods_generic("generic")
methods(generic)
```

Note the false positive: `generic.skeleton()` is not a method for our generic but an existing function in the methods package. It’s picked up because method definition relies only on a naming convention. This is another reason that you should avoid using `.` in non-method function names.

Remember that apart from methods that you’ve created, and those defined in the base package, most S3 methods will not be directly accessible. You’ll need to use `getS3method("generic", "class")` to see their source code.

```{r}
getS3method("generic", "foo")
```

## Coercion
Many S3 objects can be naturally created from an existing object through coercion. If this is the case for your class, you should provide a coercion function, an S3 generic called `as_class_name`. Base R generally does not follow this convention, which can cause problems as illustrated by `as.factor()`:

* The name is confusing, since `as.factor()` is not the factor method of the `as()` generic.
* `as.factor()` is not a generic, which means that if you create a new class that could be usefully converted to a factor, you can not extend `as.factor()`.

We can fix these issues by creating a new generic coercion function and providing it with some methods:

```{r}
as_factor <- function(x, ...) {
  UseMethod("as_factor")
}
```

Every `as_y()` generic should have a `y` method that returns its input unchanged:

```{r}
as_factor.factor <- function(x, ...) x
```

This ensures that `as_factor()` works if the input is already a factor.

Two useful methods would be for character and integer vectors.

```{r}
as_factor.character <- function(x, ...) {
  factor(x, levels = unique(x))
}
as_factor.integer <- function(x, ...) {
  factor(x, levels = as.character(unique(x)))
}
```

Typically the coercion methods will either call the constructor or the helper; pick the function that makes the code simpler. Here the helper is simplest. If you use the constructor, remember to also call the validator function.

If you think your coercion function will be frequently used, it’s worth providing a default method that gives a better error message. Default methods are called when no other method is appropriate, and are discussed in more detail in inheritance.

```{r, error=TRUE}
as_factor(1)
#> Error in UseMethod("as_factor"): no applicable method for 'as_factor' applied to an object of class "c('double', 'numeric')"

as_factor.default <- function(x, ...) {
  stop(
    "Don't know how to coerce object of class ", 
    paste(class(x), collapse = "/"), " into a factor", 
    call. = FALSE
  )
}
as_factor(1)
```

## Arguments
Methods should always have the same arguments as their generics. This is not usually enforced, but it is good practice because it will avoid confusing behaviour. If you do eventually turn your code into a package, R CMD check will enforce it, so it’s good to get into the habit now.

There is one exception to this rule: if the generic has `...`, the method must still have all the same arguments (including `...`), but can also have its own additional arguments. This allows methods to take additional arguments, which is important because you don’t know what additional arguments that a method for someone else’s class might need. The downside of using `...`, however, is that any misspelled arguments will be silently swallowed.

### Exercises
1. Read the source code for `t()` and `t.test()` and confirm that `t.test()` is an S3 generic and not an S3 method. What happens if you create an object with class test and call `t()` with it? Why?
    ```{r}
    x <- structure(1:10, class = "test")
    t(x)
    ```
**A**:
    ```{r}
    t.test
    t
    ```
if we create an object with class test, `t()`, will dispatch to `t.test()`. This simply happens, because `UseMethod()` just looks for functions named `paste0("generic", ".", c(class(x), "default"))`. So `t.test()` is erroneously treated like a method of `t()`. Since `t.test()` is a generic itself and doesn’t find a method called `t.test.test()`, it dispatches to `t.test.default()`.
    ```{r}
    getS3method("t.test", "default")(x)
    
    # specific a method t.test.test
    t.test.test <- function(x) "hi"
    t(x)
    ```

2. Carefully read the documentation for `UseMethod()` and explain why the following code returns the results that it does. What two usual rules of function evaluation does `UseMethod()` violate?
    ```{r}
    g <- function(x) {
      x <- 10
      y <- 10
      UseMethod("g")
    }
    g.default <- function(x) c(x = x, y = y)
    
    x <- 1
    y <- 1
    g(x)
    ```
**A**: First consider this where `UseMethod()` is not used
    ```{r}
    gg <- function(x) {
      x <- 10
      y <- 10
      gg.default(x)
    }
    
    gg.default <- function(x) {
      print(ls.str(environment(), all.names = TRUE))
      c(x = x, y = y) # global function environment so y = 1 when called
    }
    
    x <- 1
    y <- 1
    gg(x)
    ```
`x = 10` as this is passed to `gg.default()`, and `y = 1` as this is defined in the function environment (Global environment) of `gg.default()`.  

Now consider this from `UseMethod()` help:

... environment within which the method is evaluated by the dispatch mechanism, which is as follows:

1. Find the context for the calling function (the generic): this gives us the unevaluated arguments for the original call.

2. Evaluate the object (usually an argument) to be used for dispatch, and find a method (possibly the default method) or throw an error.

3. Create an environment for evaluating the method and insert special variables (see below) into that environment. Also copy any variables in the environment of the generic that are not formal (or actual) arguments.

4. Fix up the argument list to be the arguments of the call matched to the formals of the method.

Hence we would expect `g.default` to be constructed with `x = 1` and evaluted in a created environment with `y = 10`.

    ```{r}
    g <- function(x) {
      x <- 10
      y <- 10
      UseMethod("g")
      print("hi")
    }
    
    g.default <- function(x) {
      print(ls.str(environment(), all.names = TRUE))
      c(x = x, y = y) 
    }
    
    x <- 1
    y <- 1
    g(x)
    ```

# Method dispatch
At a high-level, S3 method dispatch is simple, and revolves around two functions, `UseMethod()` and `NextMethod()`. You’ll learn about these two functions below, and then we’ll come back to some of the additional wrinkles in dispatch details.

## UseMethod()
The purpose of `UseMethod()` is to find the appropriate method to call given a generic and a class. It does this by creating a vector of function names, `paste0("generic", ".", c(class(x), "default"))`, and looking for each method in turn. As soon as it finds a matching method, it calls it. If no matching method is found, it throws an error. To explore dispatch, we’ll use `sloop::s3_dispatch()`. You give it a call to an S3 generic, and it lists all the possible methods, noting which ones exist. For example, what happens when you try to print a `POSIXct` object?

```{r}
x <- Sys.time()
s3_dispatch(print(x))
```

`print()` will look for three possible methods, of which two exist, and one, `print.POSIXct()`, will be called. The last method is always the “default” method. This doesn’t correspond to a specific class, so is a useful catch all.

## NextMethod()
Method dispatch usually terminates as soon as a matching method is found. However, methods can explicitly choose to call the next available method using `NextMethod()`. This is useful because it allows you to rely on code that others have already written, which we’ll come back to in inheritance. Let’s make `NextMethod()` concrete with an example. Here, I define a new generic (“showoff”) with three methods. Each method signals that it’s been called, and then calls the “next” method:

```{r}
showoff <- function(x) {
  UseMethod("showoff")
}
showoff.default <- function(x) {
  message("showoff.default")
  TRUE
}
showoff.a <- function(x) {
  message("showoff.a")
  NextMethod()
}
showoff.b <- function(x) {
  message("showoff.b")
  NextMethod()
}
```

Let’s create a dummy object with classes “b” and “a”. `s3_dispatch()` shows that all three potential methods are available:

```{r}
x <- new_s3_scalar(class = c("b", "a"))
s3_dispatch(showoff(x))
```

When you call `NextMethod()` it finds and calls the next available method in the dispatch list. When we call `showoff()`, the method for `b` forwards to the method for `a`, which forwards to the default method.

```{r}
showoff(x)
```

Like `UseMethod()`, the precise semantics of `NextMethod()` are complex. It doesn’t actually work with the class attribute of the object, but instead uses a special global variable (`.Class`) to keep track of which method to call next. This means that modifying the argument that is dispatched upon has no impact, and you should avoid modifying the object that is being dispatched on.

Generally, you call `NextMethod()` without any arguments. However, if you do give arguments, they are passed on to the next method, as if they’d been supplied to the generic.

### Exercises

1. Which base generic has the greatest number of defined methods?
**A**: `print()` has clearly the most
    ```{r}
    library(methods)
    objs <- mget(ls("package:base"), inherits = TRUE)
    funs <- Filter(is.function, objs)
    generics <- Filter(function(x) ("generic" %in% pryr::ftype(x)), funs)
    
    sort(
      lengths(sapply(names(generics), function(x) methods(x), USE.NAMES = TRUE)),
      decreasing = TRUE
      )[1:3]
    ```

2. Explain what is happening in the following code.
    ```{r}
    generic2 <- function(x) UseMethod("generic2")
    generic2.a1 <- function(x) "a1"
    generic2.a2 <- function(x) "a2"
    generic2.b <- function(x) {
      class(x) <- "a1"
      NextMethod()
    }
    
    generic2(new_s3_scalar(class = c("b", "a2")))
    ```
**A**:  
* We supply an object of classes b and a2 to `generic2()`, so R will look for a method `generic2.b()`
* `generic2.b()` changes the class to a1 and then calls NextMethod()
* One could think, that R calls now `generic2.a1()`. But in fact, as mentioned in the textbook, `NextMethod()`  

doesn’t actually work with the class attribute of the object, but instead uses a special global variable (`.Class`) to keep track of which method to call next.  
    ```{r}
    generic2 <- function(x) UseMethod("generic2")
    generic2.a1 <- function(x) "a1"
    generic2.a2 <- function(x) "a2"
    generic2.b <- function(x) {
      class(x) <- "a1"
      print(.Class) # still class a2
      NextMethod()
    }
    generic2(new_s3_scalar(class = c("b", "a2")))
    
    # but this changes the class order
    generic2.b <- function(x) {
      .Class <- c("b", "a1")
      print(.Class) # still class a2
      NextMethod()
    }
    generic2(new_s3_scalar(class = c("b", "a2")))
    ```

# Inheritance
The class attribute is not limited to a single string, but can be a character vector. This, along with S3 method dispatch and `NextMethod()`, gives a surprising amount of flexibility that can be used creatively to reduce code duplication. However, this flexibility can also lead to code that is hard to understand or reason about, so you are best constraining yourself to simple styles of inheritance. Here we will focus on defining subclasses that inherit their fields, and some behaviour, from a parent class.

Subclasses use a character vector for the class attribute. There are two examples of subclasses that you might have come across in base R:

* Generalised linear models are a generalisation of linear models that allow the error term to belong to a richer set of distributions, not just the normal distribution like the linear model. This is a natural case for the use of inheritance and indeed, in R, `glm()` returns objects of class `c("glm", "lm")`.

* Ordered factors are used when the levels of a factor have some intrinsic ordering, like `c("Good", "Better", "Best")`. Ordered factors are produced by `ordered()` which returns an object with class `c("ordered", "factor")`.

You can think of the glm class “inheriting” behaviour from the lm class, and the ordered class inheriting behaviour from the factor class because of the way method dispatch works. If there is a method available for the subclass, R will use it, otherwise it will fall back to the “parent” class. For example, if you “plot” a glm object, it falls back to the lm method, but if you compute the ANOVA, it uses a glm-specific method.

```{r}
mod1 <- glm(mpg ~ wt, data = mtcars)

s3_dispatch(plot(mod1))
s3_dispatch(anova(mod1))
```


