---
title: "Subsetting"
output:
  html_document:
    number_sections: yes
    theme: sandstone
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(tidy=TRUE, results='hold', fig.align='center')
```

```{r, echo=FALSE, include=FALSE}
library(lobstr)
library(pryr)
library(knitr)
library(kableExtra)
library(magrittr)
library(dplyr)
library(tidyverse)
```

# Subsetting - multiple elements
## Atomic vectors
Let’s explore the different types of subsetting with a simple vector, `x`.

```
x <- c(2.1, 4.2, 3.3, 5.4)
```

Note that the number after the decimal point gives the original position in the vector. There are six things that you can use to subset a vector:

* **Positive integers** return elements at the specified positions.
    ```{r, results='markup'}
    x <- c(2.1, 4.2, 3.3, 5.4)
    x[c(3, 1)]
    x[order(x)]
    x[c(1, 1)] # Duplicated indices yield duplicated values
    x[c(2.1, 2.9)] # Real numbers are silently truncated to integers
    ```

* **Negative integers** omit elements at the specified positions. You can’t mix positive and negative integers in a single subset.
    ```{r, results='markup', error=TRUE}
    x[-c(3, 1)]
    x[c(-1, 2)]
    ```

* **Logical vectors** select elements where the corresponding logical value is `TRUE`.  
    ```{r, results='markup', error=TRUE}
    x[c(TRUE, TRUE, FALSE, FALSE)]
    x[x > 3]
    ```
If the logical vector is shorter than the vector being subsetted, it will be *recycled* to be the same length.
    ```{r}
    x[c(TRUE, FALSE)]
    # Equivalent to
    x[c(TRUE, FALSE, TRUE, FALSE)]
    ```
A missing value in the index always yields a missing value in the output.
    ```{r}
    x[c(TRUE, TRUE, NA, FALSE)]
    ```

* **Nothing** returns the original vector. This is not useful for vectors but is very useful for matrices, data frames, and arrays. It can also be useful in conjunction with assignment. Keeps all attributes whereas normally all attributed except `class`, `names`, `dim` and `dimnames` are dropped.
    ```{r}
    x[]
    ```

* *Character vectors* (if the vector is named) to return elements with matching names.
    ```{r, results='markup'}
    (y <- setNames(x, letters[1:4]))
    y[c("d", "c", "a")]

    # Like integer indices, you can repeat indices
    y[c("a", "a", "a")]

    # When subsetting with [ names are always matched exactly
    z <- c(abc = 1, def = 2)
    z[c("a", "d")]
    ```

## Lists
Subsetting a list works in the same way as subsetting an atomic vector. Using [ will always return a list; `[[` and `$`, as described below, let you pull out the components of the list.

```{r, results='markup'}
l <- list(1, 2)
str(l)
str(l[1])
```

## Matrices and arrays
You can subset higher-dimensional structures in three ways:

* With multiple vectors.
* With a single vector.
* With a matrix.

The most common way of subsetting matrices (2d) and arrays (>2d) is a simple generalisation of 1d subsetting: you supply a 1d index for each dimension, separated by a comma. Blank subsetting is now useful because it lets you keep all rows or all columns.

```{r, results='markup'}
a <- matrix(1:9, nrow = 3)
colnames(a) <- c("A", "B", "C")
a
a[1:2, ]
a[c(TRUE, FALSE, TRUE), c("B", "A")]
```

By default, `[` will simplify the results to the lowest possible dimensionality. See simplifying vs. preserving to learn how to avoid this.

```{r}
str(a)
class(a)
class(a[1:2, ])
class(a[1, 1])
class(a[1, 1, drop = FALSE])
dim(a[1, 1, drop = FALSE])
```

Because matrices and arrays are implemented as vectors with special attributes, you can subset them with a single vector. In that case, they will behave like a vector. Arrays in R are stored in column-major order:

```{r, results='markup'}
(vals <- outer(1:5, 1:5, FUN = "paste", sep = ","))
vals[5]
vals[c(4, 15)]
```

You can also subset higher-dimensional data structures with an integer matrix (or, if named, a character matrix). Each row in the matrix specifies the location of one value, where each column corresponds to a dimension in the array being subsetted. This means that you use a 2 column matrix to subset a matrix, a 3 column matrix to subset a 3d array, and so on. The result is a vector of values:

```{r}
select <- matrix(ncol = 2, byrow = TRUE, c(
  1, 1,
  3, 1,
  2, 4
))
vals[select]
```

## Data frames
Data frames possess the characteristics of both lists and matrices: if you subset with a single vector, they behave like lists; if you subset with two vectors, they behave like matrices. We use the following data frame for examples:

```{r, results='markup'}
df <- data.frame(x = 1:3, y = 3:1, z = letters[1:3], stringsAsFactors = FALSE)
str(df)
```

### As a list
```{r, results='markup'}
(df[1]) # subsetting a list with [ always return a list
is.data.frame(df[1]) # of 1 column - 1:3
ncol(df[1])
is.list(df[1]) # of one vector - 1:3
length(df[1]) # a list of one item - 1:3
df[[1]] # use [[ or $ to access a single item
df[1][1]
identical(df[1], df[1][1]) # TRUE since df[1] is a list of only one item
df[c("x", "y")] # select multiple columns
df[T] # T is recycled 3 times 
identical(df, df[T]) # data frame
df[] # same as df, df[T]
```

### As a matrix
```{r, results='markup'}
df[1, ]
str(df[1, ]) # still a data frame
df[, 1] # matrix subsetting simplifies by default
str(df[, 1]) # now an int vector
df[, 1, drop = FALSE]
str(df[, 1, drop = FALSE]) # still a data frame
df[1:2, 2:3]
```

### As a list and matrix
Data frames can be subsetted as a list, but using a matrix argument!
```{r, results='markup'}
idx <- matrix(c(T, F, T, F, F, T, T, T, F), ncol = 3)
idx
df[idx] # returns a vector
```
This can be used to unlist a data frame:
```{r, results='markup'}
df[matrix(T)] # T is recycled 9 times 
attributes(df[matrix(T)])
unlist(df) 
attributes(unlist(df)) # retains names
unlist(df, use.names = FALSE)
identical(df[matrix(T)], unlist(df, use.names = FALSE))
```
List-matrix subsetting does not retain factor information:
```{r, results='markup'}
df <- data.frame(x = 1:3, y = 3:1, z = letters[1:3], stringsAsFactors = TRUE)
str(df)
df[1,3]
str(df[1,3])
idx <- matrix(c(F, F, T, F, F, F, F, F, F), byrow = TRUE, nrow = 3)
idx
df[idx]
str(df[idx])

```
`unlist()` also uses the underlying integers for factors:
```{r, results='markup'}
df <- data.frame(x = 1:3, y = 3:1, z = letters[1:3], stringsAsFactors = TRUE)
str(df)
df[matrix(T)] # coerced to chr
unlist(df, use.names = FALSE) # int
```

### Lists vs. data frames
There are differeces in how vector subsetting behaves:
```{r, results='markup', error=TRUE}
df <- data.frame(x = 1:3, y = 3:1, z = letters[1:3], stringsAsFactors = FALSE)
str(df)
l <- list(x = 1:3, y = 3:1, z = letters[1:3])
str(l)

df[1, ]
l[1, ] # error

str(df[1]) # data frame
str(l[1]) # list

str(df[[1]]) # int vector
str(l[[1]]) # int vector
identical(l[[1]], df[[1]])

idx <- matrix(c(T, F, T, F, F, T, T, T, F), ncol = 3)
idx
str(df[idx]) # returns a vector
str(l[idx]) # returns list containg 1st and 3rd elements, and 3 NAs for the remaining 3 Ts

l[c(T, F, T)] # return list of 1st and 3rd elements
l[T] # recycled 3 times 
identical(l, l[T])
l[] # same as l, l[T]
l[matrix(T)] # no difference to l[T], l[], l
```

### More complicated data frames
```{r, results='markup'}
df <- data.frame(x = 1:3, y = 3:1)
df$z <- list(4:5, 4:6, 4:7)
df
str(df[matrix(T)])
str(unlist(df, use.names = FALSE, recursive = FALSE))
str(unlist(df, use.names = FALSE, recursive = TRUE))
str(df[3])
purrr::vec_depth(df[3])
str(df[[3]])
purrr::vec_depth(df[[3]])
```

## Preserving dimensionality
By default, any subsetting 2d data structures with a single number, single name, or a logical vector containing a single `TRUE` will simplify the returned output as described below. To preserve the original dimensionality, you must use `drop = FALSE`.

* For matrices and arrays, any dimensions with length 1 will be dropped:
    ```{r}
    m <- matrix(1:4, nrow = 2)
    str(m[1, ])
    str(m[1, , drop = FALSE])
    ```

* Data frames with a single column will return just that column:
    ```{r}
    df <- data.frame(a = 1:2, b = 1:2)
    str(df[, "a"])
    str(df[, "a", drop = FALSE])
    
    ```

* Factor subsetting also has a `drop` argument, but the meaning it rather different. It controls whether or not levels are preserved (not the dimensionality), and it defaults to `FALSE` (levels are preserved, not simplified by default).
    ```{r}
    z <- factor(c("a", "b"))
    z[1]
    z[1, drop = TRUE]
    ```
    ```{r, results='markup'}
    df <- data.frame(x = 1:3, y = 3:1, z = letters[1:3], stringsAsFactors = TRUE)
    f <- df[1:2, 3, drop = TRUE]
    str(f)
    is.factor(f)
    f[ , drop = TRUE]
    droplevels(f)
    ```

## Exercises
1. Fix each of the following common data frame subsetting errors:
    ```{r, results='markup', eval=FALSE}
    # mtcars[mtcars$cyl = 4, ]
    mtcars[mtcars$cyl == 4, ]

    # mtcars[-1:4, ]
    mtcars[-(1:4), ]
    
    # mtcars[mtcars$cyl <= 5]
    mtcars[mtcars$cyl <= 5, ]
    
    # mtcars[mtcars$cyl == 4 | 6, ]
    mtcars[mtcars$cyl == 4 | mtcars$cyl == 6, ]
    ```

2. Why does `x <- 1:5; x[NA]` yield five missing values?  
**A**: `NA` is of class logical so gets recycled.
    ```{r, results='markup'}
    x <- 1:5
    x[NA]
    class(NA)
    
    x[NA_integer_]
    class(NA_integer_)
    ```

3. What does `upper.tri()` return? How does subsetting a matrix with it work? Do we need any additional subsetting rules to describe its behaviour?  
    ```{r, results='markup'}
    (x <- outer(1:5, 1:5, FUN = "*"))
    upper.tri(x)
    x[upper.tri(x)]
    ```
**A**: Returns a matrix of logicals the same size of a given matrix with entries `TRUE` in the lower or upper triangle. No, see above for subsetting with logical matrices.

4. Why does `mtcars[1:20]` return an error? How does it differ from the similar `mtcars[1:20, ]`?  
**A**: `mtcars[1:20]` tries to return 20 columns, there are only 11. `mtcars[1:20, ]` returns the first 20 rows of all columns.

5. Implement your own function that extracts the diagonal entries from a matrix (it should behave like `diag(x)` where x is a matrix).  
**A**:
    ```{r, results='markup'}
    m <- matrix(1:9, ncol = 3)
    m[row(m) == col(m)]
    m <- matrix(1:10, ncol = 2)
    m
    m[row(m) == col(m)]
    m <- 4
    # convert to matrix
    m[row(as.matrix(m)) == col(as.matrix(m))]
    ```

6. What `does df[is.na(df)] <- 0` do? How does it work?  
**A**: For a data frame, `is.na()` returns a logical matrix indicating where `NA`s are. Hence `df[is.na(df)] <- 0` replaces all `NA`s with `0`s.

## Subsetting - single elements
There are two other subsetting operators: `[[` and `$`. `[[` is used for extracting single values, and `$` is a useful shorthand for `[[` combined with character subsetting. `[[` is most important working with lists because subsetting a list with `[` always returns a smaller list.

```{r}
x <- list(1:3, "a", 4:5)
str(x[1]) # list of 1 element - 1:3
str(x[[1]]) # int vector - 1:3
```

Both `[[` and `$` select a single element of the list. The main difference is that `$` does not allow computed indices, whereas `[[` does. `x$name` is equivalent to `x[["name", exact = FALSE]]`.

Note that `[[` drops names whereas `[` doesn't (but see matrix examples below which seem somewhat inconsistent).

### Named vector
```{r, results='markup'}
v <- c(a = 1, b = 2)
v
v[1] # keeps name
v[[1]] # drops name
```

### List
```{r, results='markup'}
l <- list(a = 1:2, b = 3:4)
str(l)
str(l[1]) # keeps name
str(l[[1]]) # drops name
```

### List with named vector inputs
```{r, results='markup'}
x <- c(a = 1, b = 2)
y <- c(c = 2, b = 4)
l <- list(u = x, v = y)
str(l)
str(l[1]) # keeps column and vector element names
str(l[[1]]) # drops column name, keeps vector element names
```

### Matrix
```{r, results='markup'}
m <- matrix(1:4, nrow = 2)
colnames(m) <- c("a", "b")
rownames(m) <- c("x", "y")
m

m[1] # first element of vector - drops name
m[[1]] # first element of vector - drops name

m[1, ] # keeps col names
m[, 1] # keeps row names

m[] # keeps all names

m[1, 1] # first element of vector - drops name # but see next set of examples
m[1, 1, drop = FALSE] # keeps all names - still a matrix

m[[1, 1]] # first element of vector - drops name

idx <- matrix(c(T, F, F, T), ncol = 2)
m[idx] # vector - drops names
```

The behaviour of `m[1, 1]` differs if only one of colnames or rownames has been set.

```{r, results='markup'}
m <- matrix(1:4, nrow = 2)
rownames(m) <- c("x", "y")
m[1, 1] # first element of vector - but now keeps row name

rownames(m) <- NULL
colnames(m) <- c("a", "b")
m[1, 1] # first element of vector - but now keeps col name
```

### Matrix with named vector inputs
Matrices ignore the names of vector inputs:
```{r, results='markup'}
x <- c(a = 1, b = 2, c = 3, d = 4)
m <- matrix(I(x), nrow = 2)
colnames(m) <- c("x", "y")
str(m) # I is ignored and vector element names aren't retained
```

### Data frame
In general, names are kept if the extracted part is still a data frame:
```{r, results='markup'}
df <- data.frame(a = 1:2, b = 3:4)
row.names(df) <- c("x", "y")
df

df[1] # keeps column/row names - data frame
df[[1]] # drops column/row names - int vec

df[1, ] # keeps column/row names
df[ , 1] # drops names - int vec
df[ , 1, drop = FALSE]  # keeps column/row names
df[] # keeps column/row names


df[1, 1] # drops name - int vec of 1 element
df[1, 1, drop = FALSE] # keeps column/row names - data frame

df[[1, 1]] # drops column/row names - int vec of 1 element 

idx <- matrix(c(T, F, F, T), ncol = 2)
df[idx] # drops column/row names - int vec
```

### Data frame with named vector inputs
```{r, results='markup'}
x <- c(a = 1, b = 2)
y <- c(c = 3, b = 4)
# use I to protect names on vector inputs - otherwise they are removed
df <- data.frame(x = I(x), y = I(y))
row.names(df) <- c("u", "v")
df
str(df)

df[1]
str(df[1]) # data frame - keeps column/row and vector element names

df[[1]] # vector - keeps vector element names

df[1, ]
str(df[1, ]) # keeps column/row and vector element names

df[1, 1] # vector of 1, keeps vector element name

df[1, 1, drop = FALSE]
str(df[1, 1, drop = FALSE]) # data frame, keeps column/row and vector element name

df[[1, 1]] # vector of 1, drops vector element name

idx <- matrix(c(T, F, F, T), ncol = 2)
str(df[idx]) # drops column/row names and vector element names - int vec
```

## Subsetting and assignment
All subsetting operators can be combined with assignment to modify selected values of the input vector. This is mostly useful when conditionally modifying vectors:
```{r}
df <- data.frame(a = c(1, 10, NA))
df$a[df$a < 5] <- 0
df$a
```

Subsetting with nothing can be useful in conjunction with assignment because it will preserve the original object class and structure.

```{r, results='markup'}
df <- data.frame(x = 1:3, y = 3:1, z = letters[1:3], stringsAsFactors = FALSE)
df <- map(df, as.character)
str(df)
df <- data.frame(x = 1:3, y = 3:1, z = letters[1:3], stringsAsFactors = FALSE)
df[] <- map(df, as.character)
str(df)
```

With lists, you can use `[[` + assignment + `NULL` to remove components from a list. To add a literal `NULL` to a list, use `[` and `list(NULL)`:
```{r, results='markup'}
x <- list(a = 1, b = 2)
x[["b"]] <- NULL
str(x)
y <- list(a = 1)
y["b"] <- list(NULL)
str(y)
```

When subsetting and assigning to a `NULL` object, `[[` creates a list:
```{r, results='markup'}
x <- NULL
x <- 1:2
str(x) # int vector 1:2
x <- NULL
x[1] <- 1:2 # gives warning
str(x) # int vector 1
x <- NULL
x[1:2] <- 1:2 # gives warning
str(x) # int vector 1
x <- NULL
x[[1]] <- 1:2
str(x) # list of 1 int vector - 1:2

```

## Exercises

1. How would you randomly permute the columns of a data frame? Can you simultaneously permute the rows and columns in one step?  
**A**: 
    ```{r, results='markup'}
    head(iris[sample(ncol(iris))]) # permute rows
    head(iris[sample(nrow(iris)), sample(ncol(iris))]) # permute both at the same time
    ```

2. How would you select a random sample of `m` rows from a data frame? What if the sample had to be contiguous (i.e., with an initial row, a final row, and every row in between)?   
**A**: 
    ```{r, results='markup'}
    m=4
    iris[sample(nrow(iris), m),]
    
    # Blockversion
    start <- sample(nrow(iris) - m + 1, 1)
    end <- start + m - 1
    iris[start:end, ]
    ```

3. How could you put the columns in a data frame in alphabetical order?  
**A**:
    ```{r}
    head(iris[sort(names(iris))])
    ```

